Software Frameworks for Deep Learning at Scale

### 2016-07-29 report/Report/para/0-abstract.tex
added new line to comparison chart: deep learning models
updated paper to replace CaffeOnSpark with Caffe

\begin{abstract}
The study and adoption of deep learning methods has led to significant progress in different application domains. As deep learning continues to show promise and its utilization matures, so does the infrastructure and software needed to support it. Various frameworks have been developed in recent years to facilitate both implementation and training of deep learning networks. As deep learning has also evolved to scale across multiple machines, there's a growing need for frameworks that can provide full parallel support. While deep learning frameworks restricted to running on a single machine have been studied and compared, frameworks which support parallel deep learning at scale are relatively less known and well-studied. This paper seeks to bridge that gap by surveying, summarizing, and comparing frameworks which currently support distributed execution, including but not limited to Tensorflow, CNTK, Deeplearning4j, MXNet, H2O, CaffeOnSpark, Theano, and Torch. 
\end{abstract}

>>>>

\begin{abstract}
The study and adoption of deep learning methods has led to significant progress in different application domains. As deep learning continues to show promise and its utilization matures, so does the infrastructure and software needed to support it. Various frameworks have been developed in recent years to facilitate both implementation and training of deep learning networks. As deep learning has also evolved to scale across multiple machines, there's a growing need for frameworks that can provide full parallel support. While deep learning frameworks restricted to running on a single machine have been studied and compared, frameworks which support parallel deep learning at scale are relatively less known and well-studied. This paper seeks to bridge that gap by surveying, summarizing, and comparing frameworks which currently support distributed execution, including but not limited to Tensorflow, CNTK, Deeplearning4j, MXNet, H2O, Caffe, Theano, and Torch.
\end{abstract}

### 2016-07-29 report/Report/para/1-introduction.tex
updated paper to replace CaffeOnSpark with Caffe

\section{Introduction}
Deep learning has been quite successful in improving predictive power in domains such as computer vision and natural language processing. As accuracies continue to  increase, so do the complexity of network architectures and the size of the parameter space. Google's network for unsupervised learning of image features reached a billion parameters \cite{donahue2014decaf}, and was increased to 11 billion parmeters in a separate experiment at Stanford \cite{schmidhuber2015deep}. In the NLP space, Digital Reasoning Systems trained a 160 billion parameter network \cite{trask2015modeling} fairly recently. Handling problems of this size involves looking beyond the single machine, which Google first demonstrated through its distributed DistBelief framework \cite{dean2012large}.

The goal of this paper is to survey the landscape of deep learning frameworks with full support for parallelization. Three levels of parallelization exist on the hardware level: within a GPU, between GPUs on a single node, and between nodes. Two forms of parallelism also exist on the application level: model and data parallelism. Other aspects of frameworks include release date, core language, user-facing API, computation model, communication model, programming paradigm, fault tolerance, and visualization. This choice of criteria is explained in detail in Section 2. Tensorflow, CNTK, Deeplearning4j, MXNet, H2O, CaffeOnSpark, Theano, and Torch do not necessarily encompass the entire space of frameworks for deep learning, but were selected by a combination of factors: their being open-source, level of documentation, maturity as a complete product, and level of adoption by the community. These frameworks, as well as some others not included in the Section 2 chart, are examined in detail in Section 3. Section 4 discusses finer points of parallelism and scalability in deep learning which may escape but are pertinent to the framework discussion. Section 5 offers concluding remarks.

>>>>

\section{Introduction}
Deep learning has been quite successful in improving predictive power in domains such as computer vision and natural language processing. State-of-the-art performance in computer vision is driven by the convolutional neural network model, a special kind of feed-forward deep learning model. The high-level idea is to learn images filters for extracting meaningful features and predictions. On the other hand, natural language processing has had a lot of success applying recurrent neural networks, a type of feedback model well-suited for learning order and context-sensitive sequences (as in natural languages).

As accuracies continue to increase in both domains, so do the complexity of network architectures and the size of the parameter space. Google's network for unsupervised learning of image features reached a billion parameters \cite{donahue2014decaf}, and was increased to 11 billion parmeters in a separate experiment at Stanford \cite{schmidhuber2015deep}. In the NLP space, Digital Reasoning Systems trained a 160 billion parameter network \cite{trask2015modeling} fairly recently. Handling problems of this size involves looking beyond the single machine, which Google first demonstrated through its distributed DistBelief framework \cite{dean2012large}.

The goal of this paper is to survey the landscape of deep learning frameworks with full support for parallelization. Three levels of parallelization exist on the hardware level: within a GPU, between GPUs on a single node, and between nodes. Two forms of parallelism also exist on the application level: model and data parallelism. Other aspects of frameworks include release date, core language, user-facing API, computation model, communication model, deep learning types, programming paradigm, fault tolerance, and visualization. This choice of criteria is explained in detail in Section 2. Tensorflow, CNTK, Deeplearning4j, MXNet, H2O, Caffe, Theano, and Torch do not necessarily encompass the entire space of frameworks for deep learning, but were selected by a combination of factors: their being open-source, level of documentation, maturity as a complete product, and level of adoption by the community. These frameworks, as well as some others not included in the Section 2 chart, are examined in detail in Section 3. Section 4 discusses finer points of parallelism and scalability in deep learning which may escape but are pertinent to the framework discussion. Section 5 offers concluding remarks.

### 2016-07-29 report/Report/para/2-comparison.tex
updated paper to replace CaffeOnSpark with Caffe

\section{Framework Comparison}
\begin{table*}
\small
\centering
\caption{Open-source Frameworks}
\resizebox{\textwidth}{!}{
    \begin{tabular}{|m{2cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|m{2cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|m{1.7cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|}
    \hline
    Platform & Tensorflow & CNTK & Deeplearning4j & MXNet & H2O & CaffeOnSpark & Theano & Torch \\ \hline\hline
    Release Date & 2016 & 2016 & 2015 & 2015 & 2014 & 2016 & 2010 & 2011 (deep learning) \\
    \hline
    Core Language & C++  & C++ & Java  & C++ & Java & C++, Scala & C++ & C \\
    \hline
    API & C++, Python & NDL & Java, Scala & C++, Python, R, Scala, Matlab, Javascript, Go, Julia & Java, R, Python, Scala, Javascript, web-UI & Python, Matlab, Scala & Python & Lua \\
    \hline
    Synchronization Model & Sync or async & Sync & Sync & Sync or async & Async & Sync & Async & Sync \\
    \hline
    Communication Model & Parameter server & MPI & Iterative MapReduce & Parameter server & Distributed fork-join & MPI Allreduce & N/A & N/A \\
    \hline
    Multi-GPU & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
    \hline 
    Multi-node & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \xmark & \xmark \\
    \hline
    Data Parallelism & \cmark & \cmark & \cmark & \cmark& \cmark & \cmark & \cmark & \cmark \\
    \hline
    Model Parallelism & \cmark & N/A & \xmark & \cmark & \xmark & \xmark & \cmark & \cmark \\
    \hline
    Model Parallelism & \cmark & N/A & \xmark & \cmark & \xmark & \xmark & \cmark & \cmark \\
    \hline
    Deep Learning Models & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN & DBN & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN \\
    \hline
    Programming Paradigm & Imperative & Imperative & Declarative & Both & Declarative & Declarative & Imperative & Imperative \\
    \hline
    Fault Tolerance & Checkpoint-and-recovery & Checkpoint-and-resume & Checkpoint-and-resume & Checkpoint-and-resume & N/A & N/A & Checkpoint-and-resume & Checkpoint-and-resume \\
    \hline
    Visualization & Graph (interactive), training monitoring & Graph (static) & Training monitoring & None & None & Summary Statistics & Graph (static) & Plots \\
    \hline
    \end{tabular}
}
\end{table*}

@@

>>>>

\section{Framework Comparison}
\begin{table*}
\small
\centering
\caption{Open-source Frameworks}
\resizebox{\textwidth}{!}{
    \begin{tabular}{|m{2cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|m{2cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|m{1.7cm}<{\centering}|m{1.5cm}<{\centering}|m{1.5cm}<{\centering}|}
    \hline
    Platform & Tensorflow & CNTK & Deeplearning4j & MXNet & H2O & Caffe & Theano & Torch \\ \hline\hline
    Release Date & 2016 & 2016 & 2015 & 2015 & 2014 & 2014 & 2010 & 2011 (deep learning) \\
    \hline
    Core Language & C++  & C++ & Java  & C++ & Java & C++ & C++ & C \\
    \hline
    API & C++, Python & NDL & Java, Scala & C++, Python, R, Scala, Matlab, Javascript, Go, Julia & Java, R, Python, Scala, Javascript, web-UI & Python, Matlab & Python & Lua \\
    \hline
    Synchronization Model & Sync or async & Sync & Sync & Sync or async & Async & Sync & Async & Sync \\
    \hline
    Communication Model & Parameter server & MPI & Iterative MapReduce & Parameter server & Distributed fork-join & N/A & N/A & N/A \\
    \hline
    Multi-GPU & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
    \hline 
    Multi-node & \cmark & \cmark & \cmark & \cmark & \cmark & \xmark & \xmark & \xmark \\
    \hline
    Data Parallelism & \cmark & \cmark & \cmark & \cmark& \cmark & \cmark & \cmark & \cmark \\
    \hline
    Model Parallelism & \cmark & N/A & \xmark & \cmark & \xmark & \xmark & \cmark & \cmark \\
    \hline
    Deep Learning Models & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN & DBN & DBN, CNN, RNN & DBN, CNN, RNN & DBN, CNN, RNN \\
    \hline
    Programming Paradigm & Imperative & Imperative & Declarative & Both & Declarative & Declarative & Imperative & Imperative \\
    \hline
    Fault Tolerance & Checkpoint-and-recovery & Checkpoint-and-resume & Checkpoint-and-resume & Checkpoint-and-resume & N/A & N/A & Checkpoint-and-resume & Checkpoint-and-resume \\
    \hline
    Visualization & Graph (interactive), training monitoring & Graph (static) & Training monitoring & None & None & Summary Statistics & Graph (static) & Plots \\
    \hline
    \end{tabular}
}
\end{table*}

@@

Deep learning models can be categorized into three major types: deep-belief networks (DBNs), convolutional neural networks (CNNs), and recurrent neural networks (RNNs). CNNs and RNNs were briefly described in the introduction. DBNs are less domain-specific compared to CNNs and RNNs, and could be considered a precursor to CNNs, but are fundamental nonetheless.
